Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2012-12-26 spie/_region_.tex
again text done

\message{ !name(paper.tex) !offset(106) }
\subsection{Direct Mapping}
\label{sec:directMapping}

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images, as a result of camera pose inaccuracies.

\message{ !name(paper.tex) !offset(416) }

>>>>

\message{ !name(paper.tex) !offset(76) }
\section{bla}

\label{sec:simpleTextureMapping}
\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera matrics
    $P_{1..m}$. }
  \label{fig:projection}
\end{figure}

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images to texture the target plane. Each image has a
camera matrix $P_i$ for $i=1..M$, which translates a 3D point in the
world coordinate system to a 2D point or pixel in image $i$'s
coordinates. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, such as focal length and image center, as well
as extrinsic parameters which specify the rotation and translation of
the camera's position in 3D world coordinates at the time that image
$i$ is taken. These extrinsic parameters are determined by the
backpack hardware and localization algorithms \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure} and are quite noisy.

Because our backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the texture, in order to minimize any visual discontinuities or seams
that would suggest that the plane's final texture is not composed of a
single continuous image.

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images. Given our high number of input images, it
makese sense to intelligently select only the images that match up
well, as performed in Section X, but for our solution to be robust, we
must align images directly such that we maximize the accuracy of each image.

\message{ !name(paper.tex) !offset(539) }

### 2012-12-26 spie/paper.tex
again text done

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time-consuming, and as a
result, automated 3D site modeling has garnered much interest in
recent years. CITATION

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images. Given our high number of input images, it
makese sense to intelligently select only the images that match up
well, as performed in Section X, but for our solution to be robust, we
must align images directly such that we maximize the accuracy of each image.

@@

\section{Efficient Image Alignment with Geometry Information}

In this section, we describe our method of efficient and robust image
alignment. Our approach consists of four parts. First, all images are
projected onto the surface and rotated such that detected lines within
them are parallel to lines comprising the surface's boundary and
intersection with other surfaces. Next, these projections are shifted
such that interior lines near surface intersections are matched
up. Following this, occlusion checks are performed to remove invalid
parts of each image for the target surface. These first three steps
are all image-independent and are performed in parallel. Finally, we
detect SIFT feature matches between pairs of images and solve a linear
least squares problem in 2D to maximize matches.

EVERYTHING IS IN 2D BECAUSE PITCH BECOMES WHATEVO

\subsection{Geometry-based Alignment}
\label{sec:projectionAndRotation}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we can detect
lines in the projections using Hough transforms. Experience and
intuition show that walls in indoor environments contain linear
features that are either horizontal or vertical, often corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. Though not relevant in our datasets, this can be extrapolated
such that lines are oriented with a wall's boundaries, for example in
areas with a slanted roof.

@@

To perform the rotation and alignment, we do MAGIC.  Then we indicate
that magic has been done so that in the next step stuff is good. Mark
as anchor images.

\subsection{Image Occlusion}
get from github i guess i dunno. or check avz email. maybe put in appendix

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lai1999robust, lowe1999object, mikolajczyk2005performance,
  szeliski2006image}.  SiftGPU, by PERSON CITATION, is used both for
feature detection as well as pairwise matching. These matches
determine $d^x$ and $d^y$ distances between each pair of features for
two image projections, though these distances may not always be the
same for different features. Since indoor environments often contain
repetitive features such as floor tiles or doors, we need to ensure
that SIFT-based distances are reliable. In order to mitigate the
effect of incorrect matches and outliers, the RANSAC framework
\cite{fischler1981random} is used for a robust estimate of the optimal
$d^x_{i,j}$ and $d^y_{i,j}$ distances between two images $i$ and
$j$. The RANSAC framework handles the consensus-building machinery,
and requires a fitting function and a distance function. For this
application, the fitting function simply finds the average distance
between matches in a pair of images. The distance function for a pair
of points is chosen to be the difference between those points' SIFT
match distance and the average distance computed by the fitting
function. We use a 10 pixel outlier threshold, so that SIFT matches
are labeled as outliers if their horizontal or vertical distances are
not within 10 pixels of the average distance computed by the fitting
function.

We now use the $d^x_{i,j}$ and $d^y_{i,j}$ distances between each pair
of images to refine their positions using least squares. Recall that
there are a total of $M^{2}$ possible pairs of images, though we only
generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]
If only relative distances between images are included then there is
no way to determine the absolute location of any of the images and the
matrix becomes rank deficient. From Section X, a number of images were
anchored to geometry points, and thus their locations can be used to
anchor the rest in place. In case no anchor images exist, we simply
arbitrarily pick an image. In the above matrix, the first image is set
to be such an anchor.

Finally, the $N$-dimensional observation vector $\vec{\gamma}$ is
constructed using the SIFT-based distances generated in the RANSAC
matching stage. Elements in the observation vector corresponding to
anchor images are simply their locations as determined by the original
noisy localization. Thus $\vec{\gamma}$ can be written as:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there is often a break in the chain of images,
meaning that no SIFT matches are found between one segment of the
plane and another. To solve this, we add rows to the $A$ matrix and
observations to the $\vec{\gamma}$ vector corresponding to the
original noisy distances. We then solve a weighted least squares
problem where the SIFT distances and anchor values are given a high
weight e.g. 1, while the noisy distances are given a smaller weight
e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section X results in X, which is far more seamless, but a more
intelligent image selection scheme can still be used.

@@

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section X, we saw that discontinuities occur where adjacent tiles
select different images. Though image alignment reduces or removes
many such discontinuities, there are still cases where seams are
visible, for example due to lighting differences or imprecise
matching. To reduce the cases where this happens, it makes sense to
take into account image selections made by neighboring tiles while
texture mapping a given tile. By using the same image across tile
boundaries, we can eliminate a discontinuity altogether. If this is
not possible because a tile is not visible in images used by
neighboring tiles, using similar images across tile boundaries also
leads to less noticeable discontinuities.

@@

The result of this caching approach is shown in Figure X. Seams are now greatly reduced, and LIFE IS HAPPY.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Undistorted fisheye images for vertical planes are
    tilted, but their effective camera axis is more or less normal to
    the plane. (b) The effective camera axis for ceilings is at an
    angle with respect to the plane normal. (c) Wall images are
    cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted for wall and ceiling planes in Figures
\ref{fig:projectionAngles}(a) and \ref{fig:projectionAngles}(b)
respectively. These oblique camera angles for ceilings translate into
textures that span extremely large areas once projected, as shown in
Figure \ref{fig:projectionAngles}(b). Using the tile-based texture
mapping criteria from Figure \ref{fig:scoringFunction}, such
projections have highly varying scores depending on the location of a tile on
the plane, as shown again in Figure \ref{fig:projectionAngles}(b). Thus, the
tiling approach in Section \ref{sec:simpleTextureMapping} is a
good choice for texturing floors and ceilings, as it uses the
parts of image projections that maximize resolution and accuracy for their
respective plane locations, e.g. areas near point A in Figure
\ref{fig:projectionAngles}(b), but not near point B.

@@

For wall planes however, most images are taken from close distances
and head-on angles, resulting in higher resolution fronto-parallel
projections. As a result, for each tile on a wall plane, the scoring
function of Figure \ref{fig:scoringFunction} is relatively flat with
respect to a large number of captured images, as they are all more or
less head on. Thus, the scoring function is less significant for
walls, and it is conceivable to use a different texturing strategy for
them so as to directly minimize visible seams when texturing
them. This can be done by choosing the smallest possible set of images
that (a) covers the entire plane and (b) minimizes the amount of
visible seams between them. A straightforward cost function that
accomplishes the latter is the Sum of Squared Differences (SSD) of
pixels in overlapping regions between all pairs of images. Minimizing
this cost function encourages image boundaries to occur either in
featureless areas, such as bare walls, or in areas where images match
extremely well.

@@

Given that wall-texture candidate images are taken from more or less
head-on angles, and assuming only minor rotations are made in Section
\ref{sec:projectionAndRotation}, we can reason that their projections
onto the plane are approximately rectangular. By cropping them all to
be rectangular, as shown in Figure \ref{fig:projectionAngles}(c), our
problem becomes the conceptually simpler one of filling a polygon with
rectangles, such that the sum of all costs between each pair of
rectangles is minimal. We thus also retain the advantages of working
with rectangular units, as explained in Section
\ref{sec:simpleTextureMapping}.

The location and orientation of the cameras and the fisheye lenses on
the acquisition backpack is such that images nearly always contain the
entirety of the floor to ceiling range of wall planes. Images are
therefore rarely projected with one above another when texturing wall
planes. In essence, we need only to ensure lateral coverage of wall
planes, e.g. from left to right, as our images provide full vertical
coverage themselves. We can thus construct a Directed Acyclic Graph
(DAG) from the images, with edge costs defined by the SSD cost
function, and solve a simple shortest path problem to find an optimal
subset of images with regard to the cost function \cite{dijkstra}.

@@

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a long hallway. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal cost $C_0$ to(from) all images that
meet the left(right) border of the plane.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. Since these holes are rare,
and generally fairly small, we use a greedy approach, repeatedly
filling the hole with images that result in the lowest SSD costs in a
10 px border region around the hole. This method is not as optimal as
a 2D-coverage solution would be, but it is a fast approximation, and
adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the
posters in the images, which have clear misalignment over seams in
Figure \ref{fig:compareAll}(c), but are much more aligned in Figure
\ref{fig:compareAll}(d). This is because the seam minimization
approach directly reduces the cost of each image boundary, while the
tile caching method uses a scoring function that only approximates
this effect. Furthermore, seam minimization guarantees the best
selection of images, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. The seam minimization approach
is also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images.

In the context of texturing an entire 3D planar model, we choose to
apply the seam minimization approach on walls, due to its superior
output when provided with head-on images. Floors and ceilings however,
given their many images taken at oblique angles, are textured using
the tile caching method.

% \begin{figure*}
%   \centering
%   \includegraphics[width=6in]{pipelineimages.pdf}
%   \caption{Textures corresponding to each step in our texture
%   mapping pipeline.}
%   \label{fig:pipelineimages}
% \end{figure*}

\subsection{Blending}
\label{sec:blending}
We now apply the same blending process to the two texturing methods:
tile caching or seam minimization.

Although the image alignment steps and image selection in both methods
attempt to minimize all mismatches between images, there are
occasional unavoidable discontinuities in the final texture due to
different lighting conditions or inaccuracies in planar geometry or
projection. These can however be treated and smoothed over by applying
alpha blending over image seams.  Whether the units we are blending
are rectangularly-cropped images or rectangular tiles, we can apply
the same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size.  We have found $l_2 = \frac{l_1}{2}$ to
provide enough blending without blurring features. For the seam
minimization method, we have already ensured overlap between
images. To enforce consistent blending however, we add a minimum
required overlap distance of 40 px while solving the shortest path
problem in Section \ref{sec:seamMinimization}. Additionally, if images
overlap in a region greater than the overlap distance, we only apply
blending over an area equal to the overlap distance.

After blending pixels linearly across overlapping regions using alpha
blending, our texture mapping process is complete. Figures
\ref{fig:compareAll}(e) and \ref{fig:compareAll}(f) show the blended
versions of Figures \ref{fig:compareAll}(c) and
\ref{fig:compareAll}(d) respectively. It is clear that the seam
minimization approach exhibits better alignment and fewer seams
than the tile-caching method, as Figure \ref{fig:compareAll}(f) has
the best visual quality among the textures in Figure
\ref{fig:compareAll}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time-consuming, and as a
result, automated 3D site modeling has garnered much interest in
recent years. CITATION

@@

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images. Given our high number of input images, it
makese sense to intelligently select only the images that match up
well, as performed in Section \ref{sec:imageCompositing}, but for our
solution to be more robust, we perform image alignment in order
maximize the accuracy of each image.

@@

\section{2D Image Alignment with Geometry Information}
\label{sec:2dAlignment}
In this section, we describe our method of efficient and robust image
alignment. Our approach consists of four parts. First, all images are
projected onto the surface and rotated such that detected lines within
them are parallel to lines comprising the surface's boundary and
intersection with other surfaces. Next, these projections are shifted
such that interior lines near surface intersections are matched
up. Following this, occlusion checks are performed to remove invalid
parts of each image for the target surface. These first three steps
are all image-independent and are performed in parallel. Finally, we
detect SIFT feature matches between pairs of images and solve a linear
least squares problem in 2D to maximize matches.

All of our calculations and alignments are performed in 2D, partly for
efficiency, and partly because the nature of our input data is such
that the majority of localization error is 2D, in the same plane as
the surface being projected onto. Recall that our data acquisition
system is backpack-mounted, with cameras facing to the sides of the
operator. The operator attempts to walk parallel to walls, and the
localization and model-generation algorithms that provide our input
can conform the operator's path to be straight and parallel to
detected surfaces. Since the operator walks as upright as possible,
errors in roll and yaw, as well as the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect lines
in the projections using Hough transforms. Experience and intuition
show that walls in indoor environments contain linear features that
are either horizontal or vertical, often corresponding to doors,
windows, posters, etc. Thus, when texturing walls, we rotate images in
2D such that dominant lines are made to be horizontal or
vertical. Though not relevant in our datasets, this can be
extrapolated such that lines are oriented with a wall's boundaries,
for example in areas with a slanted roof.

@@

To perform the rotation and alignment, a robust method would be to
detect all line matches and solve for an optimal homography. In nearly
all cases however, we find that a single 2D rotation is sufficient to
properly align lines in an image. We also rarely have more than one
linear geometry feature to match to. As a result, we do WHATEVER WE DO, and then for each image shifted, we indicate that they have been shifted to match geometry, which is referenced in Section \ref{sec:robustSIFTFeatureMatching}

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, nothing occurs. Where there is
a mixture of both, the rectangular region is subdivided into four, and
the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion, we
remove all erroneous texture belonging to other surfaces, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lai1999robust, lowe1999object, mikolajczyk2005performance,
  szeliski2006image}.  In our implementation, we use SiftGPU, by
PERSON CITATION, for both feature detection as well as pairwise
matching. These matches determine $d^x$ and $d^y$ distances between
each pair of features for two image projections, though these
distances may not always be the same for different features. Since
indoor environments often contain repetitive features such as floor
tiles or doors, we need to ensure that SIFT-based distances are
reliable. First, we only perform alignment on pairs of images that
overlap given the original noisy poses. Secondly, we discard feature
matches that correspond to an image distance greater than 40 pixels
from what the noisy poses estimate. In order to utilize the remaining
feature matches robustly, the RANSAC framework
\cite{fischler1981random} is used to estimate the optimal $d^x_{i,j}$
and $d^y_{i,j}$ distances between two images $i$ and $j$. The RANSAC
framework requires a fitting function and a distance function. For
this application, the fitting function simply finds the average
distance between matches in a pair of images. The distance function
for a pair of points is chosen to be the difference between those
points' SIFT match distance and the average distance computed by the
fitting function. We use a 10 pixel outlier threshold, so that SIFT
matches are labeled as outliers if their horizontal or vertical
distances are not within 10 pixels of the average distance computed by
the fitting function.

We now use the calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. Recall that there are a total of $M^{2}$
possible pairs of images, though we only generate distances between
images that overlap at SIFT feature points. Given these distances and
the original image location estimates, we can solve a least squares
problem ($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} -
\vec{\gamma}||_2^2 $) to estimate the location of the images on the
plane. The $M$-dimensional vector $\vec{\beta}$ represents the unknown
$x$ location of each image on the plane for $1 \dots M$. The optimal
$x$ and $y$ locations are obtained in the same way, so we only
consider the $x$ locations here:

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]
If only relative distances between images are included, the absolute
location of the images can not be calculated, and the matrix is rank
deficient. From Section \ref{geometryBasedAlignemtn}, a number of
images are anchored to geometry points, and thus their locations can
be used to anchor the rest in place. In case no anchor images exist,
we simply arbitrarily pick an image. In the above matrix, the first
image is set to be such an anchor, simply by placing a $1$ in its
column

The $N$-dimensional observation vector $\vec{\gamma}$ is constructed
using the SIFT-based distances generated in the RANSAC matching
stage. Elements in the observation vector corresponding to anchor
images are simply their locations as determined by the original noisy
localization. Thus $\vec{\gamma}$ can be written as:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. This solution however does not make use of our noisy camera
poses, and will fail when no SIFT matches are found between one
segment of the plane and another. To account for this, we add rows to
the $A$ matrix and observations to the $\vec{\gamma}$ vector
corresponding to the original noisy distances. We then solve a
weighted least squares problem where the SIFT distances and anchor
values are given a high weight e.g. 1, while the noisy distances are
given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:simpleTextureMapping} results in X, which has far
fewer seams, but can still be improved.

@@

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section \ref{sec:simpleTextureMapping}, we saw that
discontinuities occur where adjacent tiles are textured by different
images. Though image alignment removes many such discontinuities,
there are still cases where seams are visible due to imprecise
matching or other factors such as lighting differences. To reduce the
cases where this happens, it makes sense to take into account image
selections made by neighboring tiles while texture mapping a given
tile. By using the same image across tile boundaries, we can eliminate
a discontinuity altogether. If this is not possible because a tile is
not visible in images used by neighboring tiles, using similar images
across tile boundaries also leads to less noticeable discontinuities.

@@

The result of this caching approach is shown in Figure X, where seams
are now greatly reduced as compared to Figure X2.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Images for vertical planes are tilted, but their
    effective camera axis is more or less normal to the plane. (b) The
    effective camera axis for ceilings is at an angle with respect to
    the plane normal. (c) Wall images are cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, our data comes from a mobile backpack
system. Human operators can not carry the backpack in a perfectly
upright position and are bent forwards at 15 to 20 degrees with
respect to the vertical direction. As a result, cameras facing
sideways are head on with respect to vertical walls, while cameras
oriented towards the top or bottom of the backpack are at an angle
with respect to horizontal floors and ceilings. This is depicted in
Figures \ref{fig:projectionAngles}(a) and
\ref{fig:projectionAngles}(b). These oblique camera angles for
horizontal surfaces translate into textures that span extremely large
areas once projected, as shown in Figure
\ref{fig:projectionAngles}(b). Using the tile-based texture mapping
criteria from Figure \ref{fig:scoringFunction}, such projections have
highly varying scores depending on the location of a tile on the
plane, as shown again in Figure \ref{fig:projectionAngles}(b). Thus,
the tiling approach in this section is a
good choice for texturing floors and ceilings, as it uses the parts of
image projections that maximize resolution and accuracy for their
respective plane locations, e.g. areas near point A in Figure
\ref{fig:projectionAngles}(b), but not near point B.

@@

For vertical walls, most images are taken from close distances and
head-on angles, resulting in high resolution fronto-parallel
projections. As a result, for each tile on a wall plane, the scoring
function of Figure \ref{fig:scoringFunction} is relatively flat with
respect to candidate images, as they are all more or less head
on. Thus, the scoring function is less significant for walls, and it
is conceivable to use a different texturing strategy to directly
minimize visible seams when texturing them. This is done by choosing
the smallest possible set of images that (a) covers the entire plane
and (b) minimizes the visibility of borders between them. A
straightforward cost function that accomplishes the latter is the sum
of squared differences (SSD) of pixels in overlapping regions between
all pairs of images. Minimizing this cost function encourages image
boundaries to occur either in featureless areas, such as bare walls,
or in areas where images match extremely well.

@@

Given that wall-texture candidate images are taken from more or less
head-on angles, and assuming only minor rotations are made in Section
\ref{sec:2dAlignment}, we can reason that their projections onto the
plane are approximately rectangular. By cropping them all to be
rectangular, as shown in Figure \ref{fig:projectionAngles}(c), our
problem becomes the conceptually simpler one of filling a polygon with
rectangles, such that the sum of all costs between each pair of
rectangles is minimal. We thus also retain the advantages of working
with rectangular units, as explained in Section
\ref{sec:simpleTextureMapping}.

The operator's path, and the location and orientation of the cameras
on the acquisition backpack is such that images nearly always contain
the entirety of the floor to ceiling range of wall planes. Images
therefore rarely need to be projected with one above another when
texturing wall planes. In essence, we need only to ensure lateral
coverage of wall planes, e.g. from left to right, as our images
provide full vertical coverage themselves. We can thus construct a
Directed Acyclic Graph (DAG) from the images, with edge costs defined
by the SSD cost function, and solve a simple shortest path problem to
find an optimal subset of images with regard to the SSD cost function
\cite{dijkstra}.

@@

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a hallway wall. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal and arbitrary cost $C_0$ to(from)
all images that meet the left(right) border of the plane.

@@

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. Since these holes are rare,
and generally fairly small, we use a greedy approach, repeatedly
filling the hole with images that result in the lowest SSD costs in a
10 px border region around the hole. This method is not as optimal as
a true 2D-coverage solution would be, but it is a fast approximation,
and adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images, which
have clear misalignment over seams in Figure \ref{fig:compareAll}(c),
but are much more aligned in Figure \ref{fig:compareAll}(d). This
shortest path approach approach directly reduces the cost of each
image boundary, while the tile caching method uses a scoring function
that only approximates this effect. Furthermore, seam minimization
guarantees the best selection of images, while the sequential tile
caching method may select images early on that turn out to be poor
choices once subsequent tiles have been processed. The seam
minimization approach is also far less intensive in terms of memory
usage and runtime, both during texture generation and rendering, as it
does not require discretizing planes or images.

When texturing an entire 3D planar model, we apply the seam
minimization approach on walls, due to its superior output when
provided with head-on images. Floors and ceilings however, given their
many images taken at oblique angles, are textured using the tile
caching method.

\subsection{Blending}
\label{sec:blending}
We now apply a blending procedure to both texturing methods. Although
the image alignment steps and image selection in both methods attempt
to minimize all mismatches between images, there are occasional
unavoidable discontinuities in the final texture due to different
lighting conditions or inaccuracies in planar geometry. These can
however be treated and smoothed over by applying alpha blending over
image seams.  Whether the units we are blending are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size.  We have found $l_2 = \frac{l_1}{2}$ to
provide a balance between blending and keeping features unblurred. For
the shortest path method, we have already ensured overlap between
images. To enforce consistent blending however, we add a minimum
required overlap distance of 40 px while solving the shortest path
problem in Section \ref{sec:seamMinimization}. Additionally, if images
overlap in a region greater than the overlap distance, we only apply
blending over an area equal to the overlap distance.

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits better alignment
and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) has the best visual quality among the textures
in Figure \ref{fig:compareAll}.
